Explore expression of marker genes for tumor cell states

### 2025-01-10 analyses/cell-type-ewings/exploratory_analysis/06-tumor-cell-state-assignment.Rmd

1. The output from `aucell-singler-annotation.sh` is read in and annotations are lumped so that we can easily plot the top cell types. 
2. Clustering results from `evaluate-clusters.sh` is read in and subset to the desired parameters we would like to use for assigning clusters (leiden, modularity, resolution 0.5, nearest neighbors 20). 
3. The marker genes for tumor cell states and custom gene signatures are used to calculate the sum of all marker genes in each gene set for each library. 
4. UMAP embeddings for each library are pulled out into a data frame and combined with `SingleR` annotations, cluster assignments, and marker gene expression for all marker gene sets.

@@

# get the sum of expression of all genes for each cell state
gene_exp_df <- sce_list |>
  purrr::map(\(sce) {
    cell_states |>
      purrr::map(\(state){
        calculate_sum_markers(all_markers_df, sce, state, cell_state)
      }) |>
      purrr::reduce(dplyr::inner_join, by = "barcodes")
  }) |>
  dplyr::bind_rows(.id = "library_id")

@@

The value for each of the marker gene sets corresponds to the sum of the expression of all genes in the given gene list. 
(Note that I also looked at mean and it had similar output)

@@

```{r}
# pull out columns that have sum of marker gene expression 
marker_gene_cols <- colnames(cluster_classification_df)[stringr::str_detect(colnames(cluster_classification_df), "_sum$")]

@@

    full_celltype_heatmap(filtered_df, marker_gene_cols, "cell_state")

>>>>

1. The output from `aucell-singler-annotation.sh` is read in and annotations are lumped so that we can easily plot the top cell types. 
2. Clustering results from `evaluate-clusters.sh` is read in and subset to the desired parameters we would like to use for assigning clusters (leiden, modularity, resolution 0.5, nearest neighbors 20). 
3. The marker genes for tumor cell states and custom gene signatures are used to calculate the mean of all marker genes in each gene set for each library. 
4. UMAP embeddings for each library are pulled out into a data frame and combined with `SingleR` annotations, cluster assignments, and marker gene expression for all marker gene sets.

@@

# get the mean expression of all genes for each cell state
gene_exp_df <- sce_list |>
  purrr::map(\(sce) {
    cell_states |>
      purrr::map(\(state){
        calculate_mean_markers(all_markers_df, sce, state, cell_state)
      }) |>
      purrr::reduce(dplyr::inner_join, by = "barcodes")
  }) |>
  dplyr::bind_rows(.id = "library_id")

@@

The value for each of the marker gene sets corresponds to the mean of the expression of all genes in the given gene list.

@@

```{r}
# pull out columns that have sum of marker gene expression 
marker_gene_cols <- colnames(cluster_classification_df)[stringr::str_detect(colnames(cluster_classification_df), "_mean$")]

@@

    filtered_df <- tumor_cells_df |> 
      dplyr::filter(library_id == id)|> 
      dplyr::arrange(cell_state)

    full_celltype_heatmap(filtered_df, marker_gene_cols, "cell_state", cluster_columns = FALSE)
